---
title: "TWFE with a Continuous Treatment: Simulation Evidence"
author: Sam Harper
date: today
format:
  html:
    theme: materia
    code-fold: true
    toc: true
---

# Overview

The 'DiD revolution' of 2018–2023 established that two-way fixed effects (TWFE) regressions with **binary** staggered treatments can estimate non-convex — potentially sign-reversed — weighted averages of group-time treatment effects. Less appreciated is that similar, and in some ways worse, problems arise with **continuous** treatments, even in the simplest two-period setting with no staggering at all.

This document uses simulated data to show precisely how TWFE can go wrong when treatment is a continuous dose like ambient PM$_{2.5}$. Three scenarios build from a mundane estimand mismatch to a genuine sign reversal — all with **parallel trends holding exactly by construction**.

The takeaway is not that TWFE is broken in some exotic edge case. It is that TWFE with a continuous treatment is answering a question that most researchers do not intend to ask.

# The TWFE Estimator with Continuous Exposure

## The Differential Exposure Design

The **differential exposure design** places a policy event at time $t = 1$. Units differ in their fixed baseline exposure $D_i > 0$. Before the event all units are untreated; after it, unit $i$ is exposed at dose $D_i$. The standard TWFE regression is:

$$Y_{it} = \alpha_i + \lambda_t + \beta^{TWFE}(D_i \cdot \text{Post}_t) + \varepsilon_{it}$$

where $\alpha_i$ are unit fixed effects, $\lambda_t$ are time fixed effects, and $\text{Post}_t \in \{0, 1\}$.

The object of interest for most policy applications is the **average treatment effect on the treated**, integrated over the dose distribution:

$$\text{ATT}_\circ = \mathbb{E}[\tau(D_i)] = \mathbb{E}[Y_i(D_i) - Y_i(0)]$$

where $\tau(d) = Y_i(d) - Y_i(0)$ is the *level* effect of dose $d$ relative to no treatment.

## What TWFE Actually Estimates

In the two-period case, after removing unit and time fixed effects (the within transformation), the regressor $D_i \cdot \text{Post}_t$ becomes $(D_i - \bar{D})(\text{Post}_t - \frac{1}{2})$. The TWFE estimator reduces to the OLS slope from regressing $\Delta Y_i = Y_{i1} - Y_{i0}$ on $D_i$:

$$\hat{\beta}^{TWFE} = \frac{\text{Cov}(D_i,\, \Delta Y_i)}{\text{Var}(D_i)} = \frac{\sum_i (D_i - \bar{D})\,\Delta Y_i}{\sum_i (D_i - \bar{D})^2}$$

The implicit TWFE weight on unit $i$ is $w_i \propto (D_i - \bar{D})$. Three consequences follow immediately:

1. **Negative weights**: units with $D_i < \bar{D}$ receive negative weight — their outcome changes are *subtracted* from the estimate.
2. **Zero sum**: the weights sum to zero ($\sum_i w_i = \sum_i(D_i - \bar{D}) = 0$), so $\hat\beta^{TWFE}$ is not a positively-weighted average of unit-level effects.
3. **Implicit comparison group**: TWFE treats above-average-dose units as "treated" and below-average-dose units as the comparison group, even when every unit has $D_i > 0$.

Under parallel trends, $\mathbb{E}[\Delta Y_i | D_i = d] = \lambda + \tau(d)$, so the TWFE estimator converges to:

$$\hat{\beta}^{TWFE} \xrightarrow{p} \frac{\text{Cov}(D_i,\; \tau(D_i))}{\text{Var}(D_i)}$$

This is the slope of the dose-response function $\tau(d)$ fitted by OLS — a variance-weighted average causal response. It equals $\text{ATT}_\circ$ only when $\tau(d)$ is linear in $d$ *and* the researcher is asking about a marginal effect per unit of $D$ rather than the average total effect. For nonlinear dose-response functions or questions about average level effects, TWFE gives the wrong answer — and can produce the wrong sign.

# Simulated Data

All three simulations share the same data-generating process (DGP):

-   $N = 2{,}000$ counties, two periods ($t = 0$ pre, $t = 1$ post)
-   County-level dose $D_i$ drawn from a distribution over positive reals (no $D = 0$ units, as in ambient PM$_{2.5}$ data)
-   Unit fixed effect $\alpha_i \sim \mathcal{N}(50, 10^2)$
-   Pre-period: $Y_{i0} = \alpha_i + \varepsilon_{i0}$, $\varepsilon \sim \mathcal{N}(0, 4)$
-   Post-period: $Y_{i1} = \alpha_i + 1 + \tau(D_i) + \varepsilon_{i1}$ (common time trend = 1)
-   **Parallel trends holds exactly**: the time shift $\lambda_1 - \lambda_0 = 1$ is identical for all counties

The true dose-response $\tau(d)$ is known. We compare what TWFE estimates to the ground truth.

```{r setup, message=FALSE, warning=FALSE}
#| code-fold: true

library(tidyverse)
library(fixest)
library(patchwork)
library(kableExtra)

set.seed(42)
N <- 2000

# ---- DGP helper ----
# Builds a two-period panel under the differential exposure design.
# Parallel trends holds by construction (common time_fe for all units).
make_panel <- function(D, att_fn, time_fe = 1, sigma = 2) {
  n     <- length(D)
  tau   <- att_fn(D)
  alpha <- rnorm(n, mean = 50, sd = 10)
  # compute y outside tibble() to prevent tau column shadowing the local variable
  y_pre  <- alpha                 + rnorm(n, 0, sigma)
  y_post <- alpha + time_fe + tau + rnorm(n, 0, sigma)
  tibble(
    unit   = rep(seq_len(n), 2),
    period = rep(0:1, each = n),
    D      = rep(D, 2),
    tau    = rep(tau, 2),
    y      = c(y_pre, y_post)
  )
}

# TWFE regression: Y ~ I(D * period) | unit + period
# In two periods this recovers Cov(D, DeltaY) / Var(D)
twfe_coef <- function(panel) {
  fit <- feols(y ~ I(D * period) | unit + period,
               data = panel, warn = FALSE)
  unname(coef(fit)[1])
}

# First differences within each unit
get_deltas <- function(panel) {
  panel %>%
    select(unit, period, D, y) %>%
    pivot_wider(names_from = period, values_from = y, names_prefix = "y") %>%
    mutate(delta_y = y1 - y0)
}

theme_set(theme_classic(base_size = 11) +
  theme(plot.background = element_blank()))

d_seq <- seq(1, 25, length.out = 200)
```

# Scenario 1: Linear Dose-Response — Different Estimands

## DGP

$$D_i \sim \text{Uniform}(5,\, 20), \qquad \tau(d) = 0.5\,d$$

Each additional $\mu\text{g/m}^3$ of PM$_{2.5}$ increases the health outcome by 0.5 units. The true average ATT is:

$$\text{ATT}_\circ = \mathbb{E}[\tau(D_i)] = 0.5 \cdot \mathbb{E}[D_i] = 0.5 \times 12.5 = 6.25$$

For a *linear* dose-response, the TWFE estimator is:

$$\hat\beta^{TWFE} \xrightarrow{p} \frac{\text{Cov}(D_i,\; 0.5\,D_i)}{\text{Var}(D_i)} = \frac{0.5\;\text{Var}(D_i)}{\text{Var}(D_i)} = 0.5$$

## Results

```{r scenario1, message=FALSE, warning=FALSE}
#| fig-width: 9
#| fig-height: 4

D1   <- runif(N, 5, 20)
att1 <- function(d) 0.5 * d

p1       <- make_panel(D1, att1)
beta1    <- twfe_coef(p1)
att1_avg <- mean(att1(D1))
d1       <- get_deltas(p1)

fig1_dr <- tibble(d = d_seq, att = att1(d_seq)) %>%
  ggplot(aes(x = d, y = att)) +
  geom_line(color = "#E41A1C", linewidth = 1.2) +
  geom_hline(yintercept = att1_avg, linetype = "dotted", color = "#E41A1C",
             linewidth = 0.8) +
  annotate("text", x = 22, y = att1_avg + 0.5,
           label = sprintf("ATT° = %.2f", att1_avg),
           color = "#E41A1C", size = 3.5) +
  coord_cartesian(xlim = c(0, 25)) +
  labs(x = "Dose D (µg/m³)", y = "τ(d)",
       title = "True dose-response: τ(d) = 0.5d")

fig1_sc <- d1 %>%
  ggplot(aes(x = D, y = delta_y)) +
  geom_point(alpha = 0.2, size = 0.6, color = "grey60") +
  geom_smooth(method = "lm", color = "#E41A1C", se = FALSE, linewidth = 1.1) +
  geom_hline(yintercept = att1_avg, linetype = "dotted", color = "steelblue",
             linewidth = 0.8) +
  annotate("text", x = 20.5, y = att1_avg + 0.7,
           label = sprintf("True ATT° = %.2f", att1_avg),
           color = "steelblue", size = 3.2, hjust = 1) +
  annotate("text", x = 6, y = max(d1$delta_y) - 2,
           label = sprintf("TWFE slope = %.2f", beta1),
           color = "#E41A1C", size = 3.2, hjust = 0) +
  labs(x = "Dose D (µg/m³)", y = "ΔY (post − pre)",
       title = "TWFE regression of ΔY on D")

fig1_dr | fig1_sc
```

The TWFE slope (`r round(beta1, 2)`) is approximately 0.5 — consistent with the true dose-response slope. But this is not a good answer. The true average ATT is **`r round(att1_avg, 2)`**, which is $\mathbb{E}[D_i] \approx 12.5$ times larger.

The estimand mismatch here is not a bias in the traditional sense — TWFE is consistently estimating what it can estimate. The problem is that it estimates a *marginal causal response* (how much does one extra µg/m³ change the outcome?) when most policy questions demand an *average level effect* (how much total harm does PM$_{2.5}$ cause in this population?). These are genuinely different quantities and different policy questions, even when the dose-response is linear.

A researcher who obtains a TWFE estimate of 0.5 and interprets it as the average ATT would understate the total harm by a factor of `r round(att1_avg / beta1, 0)`.

# Scenario 2: Concave Dose-Response — Wrong Magnitude

## DGP

$$D_i \sim \text{Uniform}(5,\, 20), \qquad \tau(d) = 6\!\left(\sqrt{\frac{d}{5}} - 1\right)$$

This dose-response is zero at $d = 5$ µg/m³, positive and *concave* thereafter — diminishing marginal harm with increasing exposure. Concave dose-response functions are common in environmental health: initial exposures may cause disproportionate harm, with returns diminishing at higher concentrations.

The average causal response (ACRT) at dose $d$ is the derivative:

$$\text{ACRT}(d) = \tau'(d) = \frac{6}{10}\,\left(\frac{d}{5}\right)^{-1/2} = \frac{3}{\sqrt{5d}}$$

which is itself a decreasing function of $d$.

## Results

```{r scenario2, message=FALSE, warning=FALSE}
#| fig-width: 9
#| fig-height: 4

D2   <- runif(N, 5, 20)
att2 <- function(d) 6 * (sqrt(d / 5) - 1)

p2       <- make_panel(D2, att2)
beta2    <- twfe_coef(p2)
att2_avg <- mean(att2(D2))
d2       <- get_deltas(p2)

acrt2_med <- 3 / sqrt(5 * median(D2))  # tau'(d) at median

fig2_dr <- tibble(d = d_seq, att = att2(d_seq)) %>%
  filter(d >= 5) %>%
  ggplot(aes(x = d, y = att)) +
  geom_line(color = "#377EB8", linewidth = 1.2) +
  geom_hline(yintercept = att2_avg, linetype = "dotted", color = "#377EB8",
             linewidth = 0.8) +
  annotate("text", x = 22, y = att2_avg + 0.2,
           label = sprintf("ATT° = %.2f", att2_avg),
           color = "#377EB8", size = 3.5) +
  coord_cartesian(xlim = c(0, 25), ylim = c(-0.5, 8)) +
  labs(x = "Dose D (µg/m³)", y = "τ(d)",
       title = "True dose-response: concave")

fig2_sc <- d2 %>%
  ggplot(aes(x = D, y = delta_y)) +
  geom_point(alpha = 0.2, size = 0.6, color = "grey60") +
  geom_smooth(method = "lm", color = "#377EB8", se = FALSE, linewidth = 1.1) +
  geom_hline(yintercept = att2_avg, linetype = "dotted", color = "steelblue",
             linewidth = 0.8) +
  annotate("text", x = 20.5, y = att2_avg + 0.25,
           label = sprintf("True ATT° = %.2f", att2_avg),
           color = "steelblue", size = 3.2, hjust = 1) +
  annotate("text", x = 6, y = max(d2$delta_y) - 2,
           label = sprintf("TWFE slope = %.2f", beta2),
           color = "#377EB8", size = 3.2, hjust = 0) +
  labs(x = "Dose D (µg/m³)", y = "ΔY (post − pre)",
       title = "TWFE regression of ΔY on D")

fig2_dr | fig2_sc
```

The TWFE estimate (`r round(beta2, 2)`) is close to the true average causal response at the median dose ($d = `r round(median(D2), 1)`$ µg/m³), which is `r round(acrt2_med, 3)`. That might look like an acceptable result — but the true average ATT is **`r round(att2_avg, 2)`**, roughly `r round(att2_avg / beta2, 0)` times larger.

The reason TWFE underestimates is systematic. The TWFE variance weight $w_i \propto (D_i - \bar{D})$ is positive for high-dose units and negative for low-dose units. Because the dose-response is *concave*, the function is steepest at low doses (where units receive negative or small TWFE weight) and flattest at high doses (where units receive the largest positive TWFE weight). TWFE overweights the flat end of the curve and underweights the steep end, pulling the estimate down toward the flat region.

The practical upshot: if a researcher uses TWFE on data generated by a concave dose-response (as is typical in environmental health), the magnitude of the estimated effect will be too small regardless of sample size. This is not a finite-sample problem — it is an asymptotic bias from variance weighting.

# Scenario 3: Wrong Sign

This is the most consequential failure mode. The dose-response is *positive everywhere* — PM$_{2.5}$ causes harm at every exposure level — but TWFE returns a negative coefficient.

## DGP

$$D_i = \begin{cases} \mathcal{N}(5,\;1^2) & 50\%\text{ of counties (rural)} \\ \mathcal{N}(17,\;1^2) & 50\%\text{ of counties (industrial)} \end{cases}, \qquad \tau(d) = 20\,e^{-d/4}$$

The dose-response is steeply declining: rural counties near $d = 5$ have $\tau(5) \approx 5.7$, while industrial counties near $d = 17$ have $\tau(17) \approx 0.3$. The population average ATT $\approx 3.1$ — positive at every dose.

The bimodal dose distribution places the mean $\bar{D} \approx 11$ squarely *between* the two modes. The TWFE weight $w_i \propto (D_i - 11)$ is therefore strongly negative for rural counties (which have the largest effects) and strongly positive for industrial counties (which have the smallest effects). Positively-weighting small effects and negatively-weighting large effects produces a negative sum — even though no individual unit has a negative effect.

## Results

```{r scenario3, message=FALSE, warning=FALSE}
#| fig-width: 10
#| fig-height: 8

D3   <- c(rnorm(N / 2, mean = 5, sd = 1), rnorm(N / 2, mean = 17, sd = 1))
att3 <- function(d) 20 * exp(-d / 4)

p3       <- make_panel(D3, att3)
beta3    <- twfe_coef(p3)
att3_avg <- mean(att3(D3))
d3       <- get_deltas(p3)
D3_bar   <- mean(D3)

pA <- tibble(D = D3) %>%
  ggplot(aes(x = D)) +
  geom_histogram(bins = 60, fill = "steelblue", color = "white",
                 linewidth = 0.2) +
  geom_vline(xintercept = D3_bar, linetype = "dashed", color = "red",
             linewidth = 0.8) +
  annotate("text", x = D3_bar + 0.6, y = 75,
           label = sprintf("D̄ = %.1f", D3_bar),
           color = "red", hjust = 0, size = 3.5) +
  labs(x = "PM₂.₅ (µg/m³)", y = "Count",
       title = "A. Bimodal exposure distribution",
       subtitle = "50% rural (D ≈ 5), 50% industrial (D ≈ 17)")

pB <- tibble(d = d_seq, att = att3(d_seq)) %>%
  ggplot(aes(x = d, y = att)) +
  geom_line(color = "#377EB8", linewidth = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = att3_avg, linetype = "dotted", color = "#377EB8",
             linewidth = 0.8) +
  annotate("text", x = 20, y = att3_avg + 0.9,
           label = sprintf("ATT° = %.2f", att3_avg),
           color = "#377EB8", size = 3.5) +
  coord_cartesian(xlim = c(0, 25)) +
  labs(x = "PM₂.₅ dose (µg/m³)", y = "τ(d)",
       title = "B. True dose-response",
       subtitle = "Positive everywhere — pollution always harmful")

pC <- d3 %>%
  mutate(weight = D - D3_bar,
         sign   = if_else(weight < 0,
                          "Negative (D < D̄)", "Positive (D > D̄)")) %>%
  ggplot(aes(x = D, y = weight, color = sign)) +
  geom_point(alpha = 0.3, size = 0.7) +
  geom_hline(yintercept = 0, linewidth = 0.5) +
  geom_vline(xintercept = D3_bar, linetype = "dashed", color = "grey40") +
  scale_color_manual(
    values = c("Negative (D < D̄)" = "#E41A1C",
               "Positive (D > D̄)" = "#377EB8"),
    name = "TWFE weight"
  ) +
  labs(x = "PM₂.₅ (µg/m³)", y = "D_i − D̄",
       title = "C. TWFE weights by county",
       subtitle = "Rural counties (large τ) receive negative weight") +
  theme(legend.position = "bottom", legend.text = element_text(size = 9))

pD <- d3 %>%
  ggplot(aes(x = D, y = delta_y)) +
  geom_point(alpha = 0.2, size = 0.6, color = "grey60") +
  geom_smooth(method = "lm", color = "#E41A1C", se = FALSE, linewidth = 1.1) +
  geom_hline(yintercept = att3_avg, linetype = "dotted", color = "#377EB8",
             linewidth = 0.8) +
  annotate("text", x = 20.5, y = att3_avg + 1,
           label = sprintf("True ATT° = %.2f", att3_avg),
           color = "#377EB8", size = 3.2, hjust = 1) +
  annotate("text", x = 3, y = min(d3$delta_y) + 2,
           label = sprintf("TWFE slope = %.2f", beta3),
           color = "#E41A1C", size = 3.2, hjust = 0) +
  labs(x = "PM₂.₅ (µg/m³)", y = "ΔY (post − pre)",
       title = "D. TWFE regression of ΔY on D",
       subtitle = "Negative slope despite positive effects everywhere")

(pA | pB) / (pC | pD) +
  plot_annotation(
    title = "Scenario 3: TWFE Recovers the Wrong Sign",
    subtitle = sprintf(
      "True ATT° = +%.2f (PM₂.₅ is harmful everywhere);   TWFE = %.2f (says it's beneficial)",
      att3_avg, beta3
    )
  )
```

The TWFE estimate is **`r round(beta3, 2)`** — negative — despite every county having a positive treatment effect. A researcher relying on this result would conclude that PM$_{2.5}$ *reduces* health harm. The true average ATT is **+`r round(att3_avg, 2)`**.

This requires neither a parallel trends violation nor measurement error nor model misspecification. The DGP satisfies parallel trends exactly. The only requirements are:

1.  A dose-response that *decreases* with dose (large effects at low exposure, small at high exposure), and
2.  A dose distribution with mass concentrated at the extremes so that $\bar{D}$ lies between them.

Both are empirically plausible for ambient PM$_{2.5}$. Rural areas have lower concentrations; industrial and urban areas have higher concentrations. There are genuine scientific reasons — healthy survivor selection, compositional differences in PM$_{2.5}$ chemistry, adaptive responses — to expect attenuated health signals in chronically high-exposure populations. Whether the actual dose-response is declining over U.S. PM$_{2.5}$ ranges is an empirical question. This simulation shows that if it is, TWFE produces a qualitatively wrong answer.

## The Weight Anatomy

```{r weight_anatomy}
#| fig-width: 7
#| fig-height: 5

weight_df <- tibble(
  D      = D3,
  att    = att3(D3),
  weight = D3 - D3_bar,
  sign   = if_else(D3 < D3_bar, "Negative weight", "Positive weight")
)

weight_df %>%
  ggplot(aes(x = weight, y = att, color = sign)) +
  geom_point(alpha = 0.3, size = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(
    values = c("Negative weight" = "#E41A1C",
               "Positive weight" = "#377EB8"),
    name = ""
  ) +
  annotate("text", x = -10, y = 12,
           label = "Large τ,\nnegative weight\n(rural areas)",
           color = "#E41A1C", size = 3.5) +
  annotate("text", x = 5, y = 0.4,
           label = "Small τ,\npositive weight\n(industrial areas)",
           color = "#377EB8", size = 3.5) +
  labs(
    x = "TWFE weight (D_i − D̄)",
    y = "True τ(D_i)",
    title = "TWFE weight vs. unit's true treatment effect",
    subtitle = sprintf(
      "Correlation = %.3f — units with the largest effects are negatively weighted",
      cor(weight_df$weight, weight_df$att)
    )
  ) +
  theme(legend.position = "none")
```

The correlation between a county's TWFE weight and its true treatment effect is **`r round(cor(weight_df$weight, weight_df$att), 3)`**. The estimator is nearly perfectly designed to discount large effects and amplify small ones. This is not an unlucky simulation draw — it follows directly from variance weighting combined with a dose-response that is decreasing in $d$.

# Scenario 4: Always-On Treatment — No Post-Period

Scenarios 1–3 all assumed a clean *differential exposure design*: a policy event occurs, it "activates" a fixed county-level dose $D_i$, and we can form the $D_i \cdot \text{Post}_t$ regressor. Most ambient PM$_{2.5}$ research lacks this structure. There is no event, no "before the dose." PM$_{2.5}$ varies continuously across both units and time, and every county is exposed in every year.

The canonical always-on regression is:

$$Y_{it} = \alpha_i + \lambda_t + \beta^{TWFE} D_{it} + \varepsilon_{it}$$

where $D_{it}$ is now time-varying. After removing unit and time fixed effects, TWFE identifies off **within-unit, within-year** variation in PM$_{2.5}$: year-to-year fluctuations around each county's mean. The question is: what drives those fluctuations, and is that variation exogenous?

For ambient PM$_{2.5}$, the answer is almost certainly no. Year-to-year variation within a county reflects meteorology (cold winters increase heating demand and particle emissions), economic cycles (recessions reduce industrial output and thus PM$_{2.5}$), and wildfire activity — all of which also affect health outcomes through pathways independent of PM$_{2.5}$. TWFE cannot separate the causal PM$_{2.5}$ effect from these simultaneous influences.

## DGP

The true causal response is $\tau = 0.5$ (each µg/m³ increases the mortality rate by 0.5 units). An unobserved economic activity factor $C_{it}$ follows a persistent AR(1) process and simultaneously:

- **Drives PM$_{2.5}$ upward**: higher economic activity → more industrial output → more pollution.
- **Drives mortality downward**: higher economic activity → better incomes, healthcare access, nutrition → lower mortality, independently of pollution.

$$D_{it} = \mu_i + C_{it} + u_{it}, \qquad C_{it} = 0.9\,C_{i,t-1} + \eta_{it},\quad \eta \sim \mathcal{N}(0,\, 0.7^2)$$

$$Y_{it} = \alpha_i + \lambda_t + 0.5\,D_{it} + (-2)\,C_{it} + \varepsilon_{it}$$

$u_{it} \sim \mathcal{N}(0,\, 1.5^2)$ is exogenous weather-driven noise. The persistence $\rho = 0.9$ means the economic factor moves slowly — it dominates the *level* of PM$_{2.5}$ within a county but contributes only modestly to year-to-year *changes* (since $\text{Var}(\Delta C) = 2\sigma_\eta^2/(1+\rho) \approx 0.52$, versus $\text{Var}(\Delta u) = 2\sigma_u^2 = 4.5$).

The TWFE bias formula for this DGP is:

$$\text{plim}(\hat\beta^{TWFE}) = \tau + \delta \cdot \underbrace{\frac{\widetilde{\text{Var}}(C)}{\widetilde{\text{Var}}(C) + \widetilde{\text{Var}}(u)}}_{R^2_C \approx 0.26} \approx 0.5 + (-2)(0.26) = 0.0$$

where $\widetilde{\text{Var}}(\cdot)$ denotes within-unit, within-time variance. TWFE is predicted to give approximately zero — effectively declaring PM$_{2.5}$ to have no health effect.

## Results

```{r scenario4, message=FALSE, warning=FALSE}
#| fig-width: 10
#| fig-height: 4

set.seed(123)
N_ao <- 500; T_ao <- 10; tau_ao <- 0.5; delta_ao <- -2
rho_ao <- 0.9; s_eta <- 0.7; s_u <- 1.5

alpha_ao <- rnorm(N_ao, 50, 5)
mu_ao    <- runif(N_ao, 5, 15)
lambda_ao <- seq(0, -(T_ao - 1) * 0.3, by = -0.3)

sigma_C_ao <- s_eta / sqrt(1 - rho_ao^2)
C_mat <- matrix(NA_real_, N_ao, T_ao)
C_mat[, 1] <- rnorm(N_ao, 0, sigma_C_ao)
for (t in 2:T_ao) C_mat[, t] <- rho_ao * C_mat[, t-1] + rnorm(N_ao, 0, s_eta)

panel_ao <- expand_grid(unit = 1:N_ao, year = 1:T_ao) %>%
  mutate(
    alpha  = alpha_ao[unit],
    mu     = mu_ao[unit],
    lambda = lambda_ao[year],
    C      = C_mat[cbind(unit, year)],
    u      = rnorm(n(), 0, s_u),
    D      = mu + C + u,
    Y      = alpha + lambda + tau_ao * D + delta_ao * C + rnorm(n(), 0, 2)
  )

# Three estimators
beta_twfe_ao   <- unname(coef(feols(Y ~ D     | unit + year, panel_ao, warn = FALSE))[1])
beta_oracle_ao <- unname(coef(feols(Y ~ D + C | unit + year, panel_ao, warn = FALSE))["D"])

# Within-unit, within-time residuals for plotting
Ytilde <- resid(feols(Y ~ 0 | unit + year, panel_ao, warn = FALSE))
Dtilde <- resid(feols(D ~ 0 | unit + year, panel_ao, warn = FALSE))
Ctilde <- resid(feols(C ~ 0 | unit + year, panel_ao, warn = FALSE))

samp <- tibble(Ytilde, Dtilde, Ctilde) %>% slice_sample(n = 3000)

# Panel A: The direct confounding path: C̃ → Ỹ
pA4 <- samp %>%
  ggplot(aes(x = Ctilde, y = Ytilde)) +
  geom_point(alpha = 0.2, size = 0.5, color = "grey50") +
  geom_smooth(method = "lm", color = "#E41A1C", se = FALSE, linewidth = 1.1) +
  annotate("text", x = -3.5, y = 12,
           label = sprintf("slope ≈ %.1f\n(better economy\n→ lower mortality)", delta_ao),
           color = "#E41A1C", size = 3.2, hjust = 0) +
  labs(x = "Economic factor C̃ (within-unit demeaned)",
       y = "Mortality Ỹ",
       title = "A. Confounder acts directly on health",
       subtitle = "Better economy → more PM₂.₅ AND better health")

# Panel B: D̃ → Ỹ — what TWFE regresses
pB4 <- samp %>%
  ggplot(aes(x = Dtilde, y = Ytilde)) +
  geom_point(alpha = 0.2, size = 0.5, color = "grey50") +
  geom_smooth(method = "lm", color = "#E41A1C", se = FALSE, linewidth = 1.1) +
  geom_abline(intercept = 0, slope = tau_ao, linetype = "dashed",
              color = "#377EB8", linewidth = 1.1) +
  annotate("text", x = -5, y = 11,
           label = sprintf("TWFE slope = %.3f\n(≈ no effect)", beta_twfe_ao),
           color = "#E41A1C", size = 3.2, hjust = 0) +
  annotate("text", x = 3.5, y = -9,
           label = sprintf("True τ = %.1f", tau_ao),
           color = "#377EB8", size = 3.2, hjust = 1) +
  labs(x = "PM₂.₅ D̃ (within-unit demeaned)",
       y = "Mortality Ỹ",
       title = "B. TWFE regression of Ỹ on D̃",
       subtitle = "Confounding nearly cancels the true PM₂.₅ effect")

# Panel C: Estimator comparison
tibble(
  Estimator = factor(
    c("True τ", "Oracle TWFE\n(C known)", "TWFE\n(C omitted)"),
    levels = c("True τ", "Oracle TWFE\n(C known)", "TWFE\n(C omitted)")
  ),
  Estimate = c(tau_ao, beta_oracle_ao, beta_twfe_ao),
  Type     = c("True", "Oracle", "Biased")
) %>%
  ggplot(aes(x = Estimator, y = Estimate, fill = Type)) +
  geom_col(width = 0.5) +
  geom_hline(yintercept = tau_ao, linetype = "dashed", linewidth = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  scale_fill_manual(
    values = c("True" = "#377EB8", "Oracle" = "#4DAF4A", "Biased" = "#E41A1C"),
    guide  = "none"
  ) +
  annotate("text", x = 0.7, y = tau_ao + 0.04, label = "True τ = 0.5",
           size = 3.2, hjust = 0) +
  labs(x = NULL, y = "Estimate",
       title = "C. Estimator comparison") -> pC4

(pA4 | pB4 | pC4) +
  plot_annotation(
    title = "Scenario 4: Always-On Treatment with Time-Varying Confounding",
    subtitle = sprintf(
      "True τ = %.1f;   TWFE (no confounder control) = %.3f;   Oracle TWFE = %.3f",
      tau_ao, beta_twfe_ao, beta_oracle_ao
    )
  )
```

The TWFE estimate is **`r round(beta_twfe_ao, 3)`** — virtually zero — despite a true causal effect of 0.5. The confounding mechanism is visible in panels A and B. Panel A shows that the economic factor $C_{it}$ directly lowers mortality (better economy → better health), with a within-unit slope close to $\delta = -2$. Panel B shows that when PM$_{2.5}$ is high (positive $\tilde{D}_{it}$), mortality is also sometimes lower — because high PM$_{2.5}$ years are often high-$C$ years (economic booms drive both). The two effects nearly cancel: TWFE sees PM$_{2.5}$ increasing and mortality *not* increasing, and concludes (wrongly) that PM$_{2.5}$ has no effect.

Crucially, this failure has nothing to do with the weighting problems in Scenarios 1–3. Parallel trends is irrelevant here: the "treatment" $D_{it}$ is always on. There is no pre-period against which to validate trends, and no binary "treated/untreated" state to validate. The problem is purely time-varying confounding — the same thing that drives within-unit PM$_{2.5}$ changes also directly drives health outcomes — and TWFE has no mechanism to separate these.

# Scenario 5: The Quasi-Stayer Approach

## Operationalizing "Quasi-Stayers" for PM$_{2.5}$

De Chaisemartin, D'Haultfoeuille & Vazquez-Bare (2024) proposed an estimator for always-on continuous treatments built around the idea of **stayers** — units whose treatment doesn't change between consecutive periods. Since no county has literally the same PM$_{2.5}$ in two consecutive years, the practical extension is **quasi-stayers**: counties whose year-to-year PM$_{2.5}$ change is small enough to be treated as effectively stable.

The key idea: if a quasi-stayer's PM$_{2.5}$ barely changed, its observed health-outcome change approximates the *counterfactual trend* — what would have happened to a switcher's health had its PM$_{2.5}$ also not changed. Comparing switchers' actual outcome changes to this counterfactual, normalized by the PM$_{2.5}$ change, yields an estimate of the causal response.

**Threshold choice.** "Small" requires a threshold $\epsilon$. The cleanest approach is percentile-based: county-year $(i, t)$ is a quasi-stayer if $|\Delta D_{it}|$ falls in the bottom $q$-th percentile of the distribution of $|\Delta D|$ in year $t$. This self-scales across years regardless of units or dispersion. For PM$_{2.5}$, the bottom 20th percentile ($q = 0.20$) means counties that changed their annual mean concentration by less than roughly 0.3–0.5 µg/m³ year-to-year.

## The Estimator

For each consecutive period pair $(t-1,\, t)$:

1. **Classify** county-years: quasi-stayer if $|\Delta D_{it}| \leq \epsilon_t$; switcher otherwise.
2. **Counterfactual model**: fit $\Delta Y \sim D_{i,t-1}$ within quasi-stayers. The regressor $D_{i,t-1}$ is the *baseline dose* — conditioning on it implements de Chaisemartin's parallel trends assumption *conditional on starting dose*.
3. **Predict** each switcher's counterfactual $\widehat{\Delta Y}^{cf}$ from the quasi-stayer model evaluated at the switcher's $D_{i,t-1}$.
4. **Causal response** per switcher: $(\Delta Y_{it} - \widehat{\Delta Y}^{cf}_{it}) / \Delta D_{it}$.
5. **Aggregate** across switchers, weighting by $|\Delta D_{it}|^2$ (AOSS: average of switchers' slopes): $$\widehat{AOSS} = \frac{\sum_{i,t\,:\,\text{switcher}} (\Delta Y_{it} - \widehat{\Delta Y}^{cf}_{it})\cdot\Delta D_{it}}{\sum_{i,t\,:\,\text{switcher}} (\Delta D_{it})^2}$$

The weighting by $\Delta D_{it}^2$ gives more influence to switchers with larger dose changes — those providing the strongest signal about the slope.

## Why Quasi-Stayers Are Less Biased Here

In our DGP, PM$_{2.5}$ changes follow:

$$\Delta D_{it} = \Delta C_{it} + \Delta u_{it}$$

The exogenous noise $u_{it}$ is i.i.d., so $\text{Var}(\Delta u) = 2\sigma_u^2 = 4.5$. The economic factor is highly persistent ($\rho = 0.9$), so year-to-year changes in $C$ are small: $\text{Var}(\Delta C) = 2\sigma_\eta^2/(1+\rho) \approx 0.52$. Roughly **90% of year-to-year PM$_{2.5}$ variation is driven by the exogenous noise**, even though the economic factor dominates the *level*.

The quasi-stayer's plim has the same structural form as TWFE, but the relevant variance ratio is now over first differences rather than levels:

$$\text{plim}(\widehat{AOSS}) = \tau + \delta \cdot \underbrace{\frac{\text{Var}(\Delta C)}{\text{Var}(\Delta C) + \text{Var}(\Delta u)}}_{R^2_{qs} \approx 0.10} \approx 0.5 + (-2)(0.10) = 0.30$$

The confounded share drops from $R^2_C \approx 0.26$ (levels, used by TWFE) to $R^2_{qs} \approx 0.10$ (changes, used by quasi-stayer) — a 60% reduction in bias — because the persistent economic factor contributes far less to year-to-year fluctuations than to long-run county levels.

## Results

```{r scenario5, message=FALSE, warning=FALSE}
#| fig-width: 10
#| fig-height: 8

# First differences from panel_ao
diffs_ao <- panel_ao %>%
  arrange(unit, year) %>%
  group_by(unit) %>%
  mutate(
    dD    = D - lag(D),
    dY    = Y - lag(Y),
    dC    = C - lag(C),   # latent; used to colour points, not in estimator
    D_lag = lag(D)
  ) %>%
  filter(!is.na(dD)) %>%
  ungroup()

# Quasi-stayer estimator
qs_est <- function(df, q = 0.20) {
  df %>%
    group_by(year) %>%
    mutate(
      eps   = quantile(abs(dD), q),
      is_qs = abs(dD) <= eps
    ) %>%
    group_modify(function(d, ...) {
      qs <- filter(d,  is_qs)
      sw <- filter(d, !is_qs)
      if (nrow(qs) < 10 | nrow(sw) < 10) return(tibble(est = NA_real_))
      cf <- lm(dY ~ D_lag, data = qs)
      sw <- mutate(sw, ce = dY - predict(cf, newdata = sw))
      tibble(est = sum(sw$ce * sw$dD) / sum(sw$dD^2))
    }) %>%
    pull(est) %>%
    mean(na.rm = TRUE)
}

beta_qs <- qs_est(diffs_ao, q = 0.20)

# Mark quasi-stayer status for a single year (for plotting)
yr_plot <- 5L   # pick one representative year-pair
diffs_yr <- diffs_ao %>%
  filter(year == yr_plot) %>%
  mutate(
    eps   = quantile(abs(dD), 0.20),
    is_qs = abs(dD) <= eps
  )

# Panel A: ΔD vs ΔY scatter with QS/switcher colouring
pA5 <- diffs_yr %>%
  ggplot(aes(x = dD, y = dY, color = is_qs)) +
  geom_point(alpha = 0.35, size = 0.9) +
  # QS regression line: the counterfactual trend
  geom_smooth(data = filter(diffs_yr,  is_qs), method = "lm",
              aes(x = dD, y = dY), inherit.aes = FALSE,
              color = "grey30", se = TRUE, linewidth = 1.1,
              fill = "grey70", alpha = 0.3) +
  # TWFE slope (biased)
  geom_smooth(method = "lm", color = "#E41A1C", se = FALSE,
              linewidth = 1.1, linetype = "dashed") +
  scale_color_manual(
    values = c("TRUE" = "grey50", "FALSE" = "#377EB8"),
    labels = c("TRUE" = "Quasi-stayer (|ΔD| ≤ ε)", "FALSE" = "Switcher (|ΔD| > ε)"),
    name   = ""
  ) +
  annotate("text", x = -5, y = 14,
           label = "QS trend\n(counterfactual)",
           color = "grey30", size = 3) +
  annotate("text", x = 3, y = -10,
           label = sprintf("TWFE slope = %.3f", beta_twfe_ao),
           color = "#E41A1C", size = 3) +
  labs(x = "ΔD (PM₂.₅ change, µg/m³)",
       y = "ΔY (mortality change)",
       title = "A. QS counterfactual vs TWFE slope",
       subtitle = "Grey line = quasi-stayer counterfactual trend (near-flat)") +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 9))

# Panel B: Sensitivity to threshold q
sens_df <- tibble(
  q   = seq(0.10, 0.40, by = 0.05),
  est = map_dbl(q, ~qs_est(diffs_ao, q = .x))
)

pB5 <- sens_df %>%
  ggplot(aes(x = q, y = est)) +
  geom_line(color = "#377EB8", linewidth = 1.1) +
  geom_point(color = "#377EB8", size = 2) +
  geom_hline(yintercept = tau_ao,     linetype = "dashed", color = "grey30") +
  geom_hline(yintercept = beta_twfe_ao, linetype = "dotted", color = "#E41A1C",
             linewidth = 0.8) +
  annotate("text", x = 0.39, y = tau_ao + 0.02,
           label = "True τ = 0.5", size = 3, hjust = 1, color = "grey30") +
  annotate("text", x = 0.39, y = beta_twfe_ao - 0.03,
           label = sprintf("TWFE = %.3f", beta_twfe_ao),
           size = 3, hjust = 1, color = "#E41A1C") +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Quasi-stayer threshold (percentile of |ΔD|)",
       y = "QS estimate",
       title = "B. Sensitivity to threshold choice",
       subtitle = "Estimate stable across q — not an artefact of the threshold")

# Panel C: Side-by-side comparison of all estimators for this DGP
tibble(
  Estimator = factor(
    c("True τ", "Oracle TWFE\n(C known)", "Quasi-stayer\n(q = 0.20)", "TWFE\n(C omitted)"),
    levels = c("True τ", "Oracle TWFE\n(C known)", "Quasi-stayer\n(q = 0.20)", "TWFE\n(C omitted)")
  ),
  Estimate = c(tau_ao, beta_oracle_ao, beta_qs, beta_twfe_ao),
  Type = c("True", "Oracle", "QS", "Biased")
) %>%
  ggplot(aes(x = Estimator, y = Estimate, fill = Type)) +
  geom_col(width = 0.55) +
  geom_hline(yintercept = tau_ao, linetype = "dashed", linewidth = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  scale_fill_manual(
    values = c("True" = "#377EB8", "Oracle" = "#4DAF4A",
               "QS" = "#FF7F00", "Biased" = "#E41A1C"),
    guide = "none"
  ) +
  annotate("text", x = 0.7, y = tau_ao + 0.03,
           label = "True τ = 0.5", size = 3, hjust = 0) +
  labs(x = NULL, y = "Estimate",
       title = "C. All estimators compared") -> pC5

(pA5 | pB5 | pC5) +
  plot_annotation(
    title = "Scenario 5: Quasi-Stayer Estimator",
    subtitle = sprintf(
      "True τ = %.1f;   TWFE = %.3f;   Quasi-stayer (q=0.20) = %.3f;   Oracle = %.3f",
      tau_ao, beta_twfe_ao, beta_qs, beta_oracle_ao
    )
  )
```

The quasi-stayer estimate is **`r round(beta_qs, 3)`** — substantially closer to the truth (0.5) than TWFE (`r round(beta_twfe_ao, 3)`), though still biased. The bias ratio is about `r round(abs(beta_qs - tau_ao) / abs(beta_twfe_ao - tau_ao), 2)` — quasi-stayer retains about `r round(abs(beta_qs - tau_ao) / abs(beta_twfe_ao - tau_ao) * 100)`% of TWFE's bias. Panel A shows why: the quasi-stayer trend line (grey, estimated within counties with near-zero PM$_{2.5}$ changes) is nearly flat and provides a valid counterfactual trend for switchers; by subtracting it before dividing by $\Delta D$, the quasi-stayer estimator partially decouples the PM$_{2.5}$ effect from the economic confound.

Panel B shows that the estimate is **stable across all threshold choices** — from the 10th to 40th percentile the estimate ranges only from `r round(min(sens_df$est), 3)` to `r round(max(sens_df$est), 3)`. This stability is reassuring: the result is not an artefact of a particular cutoff.

## Limits of the Quasi-Stayer Approach

The quasi-stayer estimator reduces but does not eliminate the confounding bias. Its residual bias ($\approx$ `r round(abs(beta_qs - tau_ao), 2)`) reflects the share of year-to-year PM$_{2.5}$ variation that is still driven by the economic factor ($R^2_{qs} \approx 0.10$). More importantly, the identifying assumption is:

> *Counties with similar baseline PM$_{2.5}$ whose annual concentration barely changed would have experienced the same mortality trend as counties whose concentration changed substantially, had those counties also not changed.*

This fails whenever the factor that caused the PM$_{2.5}$ change is itself directly affecting health. In our DGP the economic factor drives both PM$_{2.5}$ changes and health, so the assumption is technically violated — only the high persistence of $C$ (which reduces $R^2_{qs}$ relative to $R^2_C$) limits the damage. In settings where PM$_{2.5}$ changes are driven primarily by transient weather shocks (which also directly affect health), the quasi-stayer may perform no better than TWFE.

The assumption is, however, **partially testable**: using a period earlier than $t-1$ as a placebo, one can check whether quasi-stayers and switchers had differential outcome trends even in a period when PM$_{2.5}$ did not change. If the identifying assumption holds, the quasi-stayer estimator should return approximately zero for placebo treatment changes.

# Summary

```{r summary_table}
tibble(
  Scenario        = c("1: Linear τ(d) = 0.5d",
                      "2: Concave τ(d) = 6(√(d/5)−1)",
                      "3: Declining τ(d) = 20exp(−d/4)",
                      "4: Always-on + econ. confounder",
                      "5: Quasi-stayer (q = 0.20)"),
  Setting         = c(rep("Differential exposure (Post exists)", 3),
                      rep("Always-on (no Post)", 2)),
  `PT holds?`     = c("Yes", "Yes", "Yes", "No", "Partially"),
  `TWFE`          = c(round(beta1, 2), round(beta2, 2), round(beta3, 2),
                      round(beta_twfe_ao, 2), round(beta_twfe_ao, 2)),
  `Alt. estimate` = c(rep("—", 3), round(beta_oracle_ao, 2), round(beta_qs, 2)),
  `True`          = c(round(att1_avg, 2), round(att2_avg, 2), round(att3_avg, 2),
                      tau_ao, tau_ao),
  `Failure mode`  = c("Slope ≠ level effect",
                      "Variance-weighting underestimates concave ATT",
                      "Wrong sign (declining ATT × bimodal D)",
                      "Time-varying confounding cancels true effect",
                      "Bias reduced ~60% vs TWFE; not eliminated")
) %>%
  kbl(align = "lllrrrl",
      col.names = c("Scenario", "Setting", "PT holds?", "TWFE",
                    "Alt.", "True", "Failure mode")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  row_spec(3, bold = TRUE, color = "white", background = "#C0392B") %>%
  row_spec(4, bold = TRUE, color = "white", background = "#8B0000") %>%
  row_spec(5, bold = TRUE, background = "#FFF3CD")
```

## What TWFE Is (and Isn't) Doing

Across Scenarios 1–3, parallel trends holds by construction. TWFE fails because of *what it estimates* rather than because of assumption violations. In the two-period differential exposure case:

$$\hat\beta^{TWFE} \xrightarrow{p} \underbrace{\frac{\text{Cov}(D_i,\; \tau(D_i))}{\text{Var}(D_i)}}_{\text{what TWFE estimates}} \quad \neq \quad \underbrace{\mathbb{E}[\tau(D_i)]}_{\text{average ATT}}$$

The numerator is the covariance between dose and dose-response — a slope, not a level. These two coincide only when $\tau(d)$ is linear in $d$, and even then TWFE recovers the marginal causal response per unit of dose rather than the average level effect. When the dose-response is decreasing in dose (Scenario 3), the covariance is negative and TWFE reverses sign.

In Scenarios 4–5 the failure mode is different: the "treatment" is always on, there is no untreated state, and the within-unit variation that TWFE exploits is contaminated by time-varying confounders. Parallel trends cannot be invoked or tested in the usual way, because there is no pre-treatment period to validate.

Note that the Scenarios 1–3 pathologies arise with just **two time periods** — unlike staggered binary treatment problems, which require multiple periods. The problems are about the continuous treatment distribution and dose-response shape, not timing.

## Implications

**First**, TWFE with a continuous treatment in the differential exposure design implicitly runs a binary DiD: above-average-dose units as "treated," below-average-dose as "control." Whether this is scientifically appropriate depends on the dose-response shape — and there is no test that can reveal this without knowing the truth.

**Second**, in the always-on setting, TWFE uses within-unit year-to-year variation in PM$_{2.5}$. Because this variation is driven partly by the same factors (weather, economic cycles) that also affect health outcomes, the identifying assumption is almost certainly violated. Pre-trend tests are not available to diagnose this.

**Third**, the quasi-stayer approach exploits a better-justified source of variation — counties with near-zero PM$_{2.5}$ changes as controls — and can reduce bias substantially when the confounder is persistent (moves slowly relative to year-to-year changes). But it is not a magic fix: if the confounder drives year-to-year PM$_{2.5}$ changes, quasi-stayer shares TWFE's problem. The assumption is more transparent and partially testable via placebo pre-period regressions.

**The honest bottom line**: for ambient PM$_{2.5}$ health effect estimation, instrumental variable methods — wind direction, wildfire smoke plumes, regulatory nonattainment thresholds — remain the most defensible identification strategy. DiD-type methods (both standard TWFE and quasi-stayer variants) can be made to work under carefully defended parallel trends or quasi-stayer assumptions, but neither assumption is likely to hold without question for a pollutant whose variation is so tightly linked to the same economic and meteorological forces that determine population health.
